\SW: II. New lens candidates from the CFHTLS discovered through citizen science

### 2015-04-10 doc/sw-cfhtls.tex
went through Sect 1-2, changed last fig.

%%% ENSURE WORD LIMIT IS 250
We report the discovery of 28 promising and 58 total new lens candidates
in the CFHT Legacy Survey data (\cfhtls) based on about 11 million
classifications performed by citizen scientists as part of the first
blind lens search undertaken by \sw.  The principle goal of this lens
search was to identify lenses missed by lens finding robots (the \rf on
galaxy scales and \af on group/cluster scales), which have been
previously used to mine the \cfhtls data for lenses.  We compare the
properties of known lens samples detected by these algorithms to the \sw
sample and found that they are broadly similar.  The image separation
distribution calculated from the combined sample of known and new lens
candidates shows that our previous constraints on the average density
profile of the lens population are robust.  In addition to recovering
about 60\% of the known sample, the new \sw candidates possess a richer
variety compared to the lenses found by the two robots. We find that
analyzing only those classifications which are performed by the more
skilled volunteers, \sw can achieve a detection rate of up to 75\% for
the known lens sample.  Training and calibration of the performance of
citizen scientists is crucial for the success of \sw. We also present
the SIMCT pipeline, used for generating a sample of realistic simulated
lensed images in the CFHTLS, and a sample of duds and false positives
used in the training.  Such a training sample has a legacy value for
testing future lens finding algorithms. We make our training sample
publicly available.

@@

%{\it Describe various lens finding algorithms and the motivation for this particular project with \cfhtls. Already searched by robots: enables comparison of techniques. Lenses not yet found by robots, detectable by humans? }

@@

The search for gravitational lenses is a needle-in-a-haystack problem.
Efficient automated methods are imperative to find a reasonably complete
and pure sample of strong lenses. Several lens finding algorithms have
been developed so far for this purpose
\citep[e.g.][]{Lenzen2004,Alard2006,Seidel2007,More2012,Gavazzi2014} but they can not
simultaneously capture the myriad types of lenses that are known to
exist. For example, the lensed images of background galaxies show
variety in their surface brightness distributions, colors, light
profiles, shapes, structures and angular image separations.  Moreover,
many lensed images appear similar to features found commonly in galaxies
(such as spiral arms) or to artifacts in astronomical images (scattered
light around stars).  Almost all lens finding algorithms find it
difficult to distinguish these from the real lenses and thus, suffer
from a high rate of false positives.  To minimize this problem,
algorithms are often restricted to detect a very narrow class of lens
systems. However, even after such restrictions, robotic lens searches
have to always rely on visual screening to arrive at a list of plausible
candidates.

Recognising patterns amidst an array of noisy pixelated images is one of the
strengths of human brains. Humans are also capable of dealing with
multi-tier complex web of questions before arriving at a conclusion which may
not be always possible to automate. The algorithm by which our brains process a
task is extremely malleable, self-learning and self-evolving. Therefore it has
a huge potential for the discovery of exotic objects which do not quite fit a
set criteria, but are still very likely to be objects of interest. The lens
finding algorithms are not yet advanced enough to produce better performance
than visual classifications. Given this, participation of a large community of
volunteers to help with the visual identification of lenses is extremely
beneficial for the lensing community. In particular, as we enter the era of
large area imaging surveys (spanning thousands of square degrees), it is
the perfect time to tap into the potential of citizen science.

\GZ was the first citizen science project in astronomy, and addressed the
problem of how to  classify large numbers of galaxies by their morphology
\citep{Lintott2008}.  From the early results based on morphological
classifications of galaxies, to several new unexpected and interesting
discoveries, such as that of green pea galaxies and Hanny's Voorwerp, \GZ
has been able to harness the potential of citizen scientists.  Since then
astronomy and non-astronomy projects have been launched under the citizen
science web portal Zooniverse (http://zooniverse.org). The task of finding
gravitational lenses is significantly challenging given the fact that these
lenses are rare. To add to the challenge, not many citizen scientists are
expected to be aware of the phenomenon of gravitational lensing, and the
resulting image configurations.  With these significant challenges at hand, we
designed the \sw project to enable the discovery of lenses through citizen
science.  In Paper I \citep{Marshall2015}, we describe the design of \sw and how the entire
system functions as a discovery service.

%
%\subsection{Paper I: Overview}
%In Paper I, we give details of the classification interface used by the
%volunteers. The regular images called test subjects are served randomly from the
%database. Some of the test subjects are labelled as training subjects because
%either these have simulated lensed images superposed around real galaxies or
%there are some false positives in the subjects.  The sample of training subjects
%is much smaller compared to the sample of test subjects. The training subjects
%are shown with decreasing frequency to the volunteers depending on how new the
%volunteers are to the project based on their classification history. However,
%the frequency of training subjects becomes a constant after XXX classifications
%from a volunteer.
%
%Clicking (or failing to click) on the training subjects results in a
%pop-up feedback message that congratulates the volunteers or points them
%to read the Spotter's guide pages. The website has several examples of
%real lenses and false positives with explanation on the Spotter's guide
%page. A quick reference to the Spotter's guide is available in the
%classification interface as well. Additional features to change zoom,
%contrast and brightness are made available to make a better judgement of
%the difficult lens candidates.
%
%The classification of subjects happens in two serial stages. \StageOne is
%for quick rejection of all uninteresting subjects. Subjects that passed
%\StageOne are then reinspected at \StageTwo, for a careful selection of
%most promising lens candidates. In general, subjects are assigned a
%small prior probability for it to contain a lens. As each subject is
%seen by multiple volunteers, the posterior probability P is updated by
%using the weights of the volunteers. The weights are determined based on
%how well the volunteers perform on the training sample. Once a subject
%crosses a threshold (acceptance or rejection), it is retired from the
%database to ensure that rest of the subjects receive sufficient multiple
%classifications. This is true for \StageOne with ~10$^5$ subjects but is
%not required for \StageTwo with only a few thousand subjects.
%
%\question{AM}{PJM,AV, should we have this overview section on paper I?}
%

In this paper ( Paper II hereafter), we describe our first lens search using
data from the Canada-France-Hawaii Telescope Legacy Survey
\citep[CFHTLS,][]{Gwyn2012}. This
paper is organised as follows. In \Sref{sec:data}, we introduce the \cfhtls
imaging data and the known lens samples from the \cfhtls. In \Sref{sec:ts} and
\Sref{sec:dfp}, we explain the training sample consisting of simulated lenses,
duds and false positives generated to aid the \sw volunteers in the lens search.
In \Sref{sec:results}, we present the new lens candidates from \sw and compare
\sw to two lens finding algorithms which produced the known \cfhtls lens
samples. Next, we discuss what kind of lenses are detected or missed by the
algorithms and \sw in \Sref{sec:discuss}. The summary and conclusions are given
in \Sref{sec:conclude}.

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec. The
\cfhtls \textsc{WIDE} covers a total non-overlapping area of 150~deg$^2$
on the sky and consists of four fields W1, W2, W3 and W4. The field W1
has the largest sky coverage of 63.65~deg$^2$. The fields W2 and W4 have
similar sky coverages of 20.32~deg$^2$ and 20.02~deg$^2$,
respectively\footnote{These numbers are estimated from
http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.  The field W3 has a
sky coverage of 42.87~deg$^2$ and is more than twice as large as W2 and
W4.

The \cfhtls imaging is very homogeneous and has good image quality. Most of
the lensed arcs are much brighter in the $g$ band thus, deep imaging in
this band is desirable. The limiting magnitude is 25.47 for the $g$ band
which goes the deepest among all of the five bands. The mean seeing in
the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the data from the
final T0007 release taken from the Terapix
website\footnote{{http://terapix.iap.fr/cplt/T0006-doc.pdf}}
for this work.

We note that the \cfhtls is a niche survey with a unique combination of wide
imaging with deep sensitivity. It is a precursor to the ongoing wide imaging
surveys such as the Dark Energy Survey (DES), Kilo Degree Survey (KiDS) and the
Hyper Suprime-Cam (HSC) survey and other planned future surveys such as the
Large Synoptic Survey Telescope (LSST) survey.  The search for lenses with \sw
in the \cfhtls is an important step to learn lessons and prepare for lens
searched in these larger imaging surveys.

@@

The \cfhtls has been searched for lenses using various lens finding
methods and algorithms. Here, we give a brief summary of previously published lens
samples in chronological order.

@@

\citet{Sygnet2010} carried out a search for edge-on galaxy lenses in the
\cfhtls WIDE. They identified galaxies, using {\sc Sextractor}, which
have $18<i<21$ and only those with inclination angle $<25$~deg. With a
few more selection criteria and visual inspection, they found about 3
promising and a total of 18 lens candidates.

The \af \citep{More2012} was used for finding blue arc-like features in
the complete \cfhtls data without any pre-selection on the type of the lensing
object. The algorithm measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales.  In the absence of a large systematically followed up
verification of the candidates, we use the  most promising sample of 26 lens
candidates as the bonafide known lens sample. And, the sample of 108 candidates
form the known lens candidates which come from the \cfhtls WIDE alone since the
\sw lens search is carried out on the WIDE imaging only.

In \citet{Elyiv2013}, they visually inspected a sample of 5500 optical
counterparts of X-ray point-like sources identified in the XMM-LSS
imaging of the \cfhtls W1 field. The goal was to discover lensed
quasars. They find 3 very promising and a total of 18 candidates.

The \rf \citep{Gavazzi2014} was used for finding compact rings or arcs
around centers of isolated and massive early-type galaxies. The \rf
subtracts the PSF-matched $i$-band images from the $g$-band images, and
looks for excess flux in the bluer $g$-band. An object detector measures
the properties of these residual blue features, and candidates which
meet the length-width ratio and tangential alignment criteria are then
visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T06 CFHTLS data release catalogs.  14370 were
found to show detectable blue residuals, and 2524 were visually
inspected, having passed the automatic feature selection process. This
led to a total of 330 lens candidates out of which 42 are good quality
(\texttt{q\_flag} = 3) and 288 are medium quality (\texttt{q\_flag} = 2)
candidates. In addition to the main well-defined sample,
\citet{Gavazzi2014} reported a further 71 candidates were reported
detected by earlier version of the \rf or from the \cfhtls DEEP.  From
the main sample of ``\rf candidates,'' the SL2S team found, during their
follow-up campaign, 33 confirmed lenses (and 17 promising candidates).

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of only 37~sq.~deg
and the entire data was visually inspected to estimate completeness and
purity of their robotic search. They found 29 candidates with the
algorithm and 41 candidates through visual inspection alone some of
which are known from previous searches. Most of these candidates are low
probability\footnote{http://www.ita.uni-heidelberg.de/˜maturi/Public/arcs}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm that has been run on the entire \cfhtls imaging.
Thus, these can be considered as the known reference sample from robotic
searches.  For the purposes of transparency and to help with the
training, the volunteers participating in \sw-\cfhtls lens search were
made aware of these two known lens samples. Images containing the
systems from the \rf and the \af samples were labelled as ``Known Lens
Candidates'' in the Talk forum, where volunteers have the opportunity to
discuss their findings with fellow volunteers and the science team. In
this paper, we refer to the sample of 330 \rf and 108 \af lens
candidates as the sample of ``known lens candidates'' and the sample of
confirmed (or most promising) 33 \rf and 26 \af as the sample of ``known
lenses''. Note that the ``known lens'' sample is a subset of the ``known lens
candidates'' sample.

@@

In order to perform a blind lens search over the entire \cfhtls
\textsc{wide}, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in Paper I. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, HumVI\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. It is possible that a lens
candidate happens to be too close to the edge of a cutout, the overlap
between neighbouring cutout allows the inspector to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, an inspector may not necessarily come
across the neighbouring cutout unless the inspector classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.
%We also provided access to the $u$ and $z$ band imaging
%data through dashboard, when the volunteers wished to take a closer look at a particular system\footnote{The dashboard tool page can be accessed from
%http://tools.zooniverse.org/\#/dashboards/spacewarp}.

@@

In this section, we describe the framework to generate the simulated lens sample
and give details of the sample itself along with some known limitations. This
simulated lens sample is an important component for analyzing the
classifications performed by the citizen scientists.

@@

For the purpose of generating simulated gravitational systems, we divide the
lenses in to two main categories a) galaxy scale lenses b) and group or
cluster-scale lenses. We further subdivide galaxy scale lenses based on the
type of the background sources, namely galaxies and quasars. We do not simulate
group scale quasar lenses as they are expected to be extremely rare. We now
describe our procedure to generate these different types of lens systems.

@@

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2005} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton:2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

@@

For quasars, we assume the luminosity function prescription of \citep{Ogur2010}
and adopt k-corrections by \citep{Richards2006}.
%alp=-0.5;
%kcorr=-2.5*(1 + alp)*log10(1+zz);
%Dlum=cc.Dlofz(zz)/p.hval;
%DM=5*log10(Dlum);
%Mabs=mag-kcorr-DM-25.0;
The luminosity function is expressed as
\be
\frac{{\rm d}\Phi}{{\rm d}M}=\frac{\Phi_{*}}{10^{0.4(\alpha+1)(M_{\rm abs}-M_{*})} + 10^{0.4(\beta+1)(M_{\rm abs}-M_{*})} }
\ee
where the normalization, $\phi_{*}=5.34\times10^{-6} h^3$ Mpc$^{-3}$ and break
magnitude, $M_*=-20.90 + 5 {\rm log} h - 2.5 {\rm log} f(z)$. The redshift
dependent factor in $M_*$ is given by
\be
f(z)=\frac{e^{\zeta z_s}(1+e^{\xi z_*})}{(\sqrt{e^{\xi z_s}}+\sqrt{e^{\xi z_*}})^2} \,.
\ee
We adopt the best-fit values $\zeta=2.98$, $\xi=4.05$, $z_{*}=1.60$
\citep{Oguri2010}. For the faint end slope, we use $\beta=-1.45$ whereas for
the bright end slope, we use $\alpha=-3.31$ when $z_s<3$ and $\alpha=-2.58$ at
higher redshifts, as prescribed by \citep{Oguri2010}.
respectively.

@@

Next, we determine properties of the background source for every lens. We follow
similar procedures for both background galaxies and quasars. We perform
ray-tracing for all of the $N_{\rm src}$ sources using the publicly available
code {\sc gravlens} \citep{Keeton2000} and choose sources that satisfy our
selection criteria. We determine fluxes of the lensed images and the total
magnification of each of the lensed source. We draw a random source for which
the flux of the second brightest lensed image and the total magnification of all
lensed images are above the thresholds given in Table~\ref{tab:thresh}.

>>>>

%%% ENSURE WORD LIMIT IS 250
We report the discovery of 28 promising and a total of 58 new lens
candidates from the CFHT Legacy Survey (\cfhtls) based on about 11
million classifications performed by citizen scientists as part of the
first \sw lens search.  The goal of the blind lens search was to
identify lenses missed by lens finding robots (the \rf on galaxy scales
and \af on group/cluster scales), which have been previously used to
mine the \cfhtls for lenses.  We compare some properties of lens samples
detected by these algorithms to the \sw sample and found that they are
broadly similar.  The image separation distribution calculated from the
\sw discovered sample shows that our previous constraints on the average
density profile of the lens population are robust. \sw recovers about
60\% of the known sample and the new candidates show a richer variety
compared to the lenses found by the two robots. We find that analyzing
only those classifications which are performed by the high power
volunteers, \sw can achieve a detection rate of up to 75\% of the known
lens sample.  Training and calibration of the performance of citizen
scientists is crucial for the success of \sw. We also present the SIMCT
pipeline, used for generating a sample of realistic simulated lensed
images in the CFHTLS, and a sample of duds and false positives used in
the training.  Such a training sample has a legacy value for testing
future lens finding algorithms. We make our training sample publicly
available.
%%% ENSURE WORD LIMIT IS 250

@@

@@

The search for gravitational lenses is a needle-in-a-haystack problem.
Efficient automated methods are imperative to find a reasonably complete
and pure sample of strong lenses. Several lens finding algorithms have
been developed so far 
\citep[e.g.][]{Lenzen2004,Alard2006,Seidel2007,More2012,Gavazzi2014} but they can not
simultaneously capture the myriad types of lenses that are known to
exist. For example, the lensed images of background galaxies show
variety in their surface brightness distributions, colors, light
profiles, shapes, structures and angular image separations.  Moreover,
many lensed images appear similar to features found commonly in galaxies
(such as spiral arms) or to artifacts in astronomical images (scattered
light around stars).  Almost all lens finding algorithms find it
difficult to distinguish these from the real lenses and thus, suffer
from the problems of high rate of false positives.  To minimize this problem,
algorithms are often restricted to detect a very narrow class of lens
systems. However, even after such restrictions, robotic lens searches
have to always rely on visual screening to find a sample of plausible
lens candidates.

Recognising patterns amidst an array of noisy pixelated images is one of the
strengths of human brains. Humans are also capable of dealing with
multi-tiered complex web of questions before arriving at a conclusion which may
not be always possible to automate. The algorithm by which our brains process a
task is extremely malleable, self-learning and self-evolving. Therefore it has
a huge potential for the discovery of exotic objects which do not quite fit a
set criteria, but are still very likely to be objects of interest. The lens
finding algorithms are not yet advanced enough to produce better performance
than visual classifications. Therefore, participation of a large community of
volunteers to help with the visual identification of lenses is extremely
beneficial for the lensing community. In particular, as we enter the era of
large area imaging surveys (spanning thousands of square degrees), it is
a perfect time to tap into the potential of citizen science.

\GZ was the first citizen science project in astronomy, and addressed
the problem of how to  classify large numbers of galaxies by their
morphology \citep{Lintott2008}.  From the early results based on
morphological classifications of galaxies, to several new unexpected and
interesting discoveries, such as that of green pea galaxies
\citep{Cardamone2009,Jaskot2013} and Hanny's Voorwerp
\citep{Lintott2009,Keel2012}, \GZ has been able to harness the potential
of citizen scientists.  Since then astronomy and non-astronomy projects
have been launched under the citizen science web portal Zooniverse
(http://zooniverse.org). The task of finding gravitational lenses is
significantly challenging given that the lens systems show such
complexity and that they are rare. To add to the challenge, not many
citizen scientists are expected to be aware of the phenomenon of
gravitational lensing, and the resulting image configurations.  With
these significant challenges at hand, we designed the \sw project to
enable the discovery of lenses through citizen science.  In a companion
paper, Paper I (Marshall et al., in prep.), we describe the design of
\sw and how the entire system functions as a discovery service.

In this paper (Paper II hereafter), we describe our first lens search
using data from the Canada-France-Hawaii Telescope Legacy Survey
(\cfhtls). This paper is organised as follows. In
\Sref{sec:data}, we introduce the \cfhtls imaging data and the
previously published lens samples from the \cfhtls. We generated a training sample, 
consisting of simulated lenses, duds and false positives, in order to
aid the \sw volunteers in the process of finding lenses. We give details of this
training sample in \Sref{sec:ts} and \Sref{sec:dfp}. In \Sref{sec:swap},
we briefly describe how the classifications of images from the volunteers are
turned into a catalog of plausible candidates (for further details, see
Paper I). In \Sref{sec:results}, we present the new lens candidates from \sw and
compare it to the lens samples produced by a few robotic searches from
the \cfhtls in the past. Next, we discuss what kind of lenses are detected
or missed by the algorithms and \sw in \Sref{sec:discuss}. The summary
and conclusions are given in \Sref{sec:conclude}.

@@

The \cfhtls is a photometric survey in five optical bands
($u^*g'r'i'z'$) carried out with the wide-field imager MegaPrime which
has a 1~deg$^2$ field-of-view and a pixel size of 0.186\arcsec
\citep{Gwyn2012}. The \cfhtls \textsc{WIDE} covers a total
non-overlapping area of 150~deg$^2$ on the sky and consists of four
fields W1, W2, W3 and W4. The field W1 has the largest sky coverage of
63.65~deg$^2$. The fields W2 and W4 have similar sky coverages of
20.32~deg$^2$ and 20.02~deg$^2$, respectively\footnote{These numbers are
estimated from http://terapix.iap.fr/cplt/table\_syn\_T0006.html}.  The
field W3 has a sky coverage of 42.87~deg$^2$ and is more than twice as
large as W2 and W4.

The \cfhtls imaging is very homogeneous and has good image quality. Most
of the lensed arcs are much brighter in the $g$ band thus, deep imaging
in this band is desirable. The limiting magnitude is 25.47 for the $g$
band which goes the deepest among all of the five bands. The mean seeing
in the $g$ band is 0.78\arcsec. The zero point to convert flux to AB
magnitude for all bands is 30. These characteristics make \cfhtls ideal
to do visual inspection for finding lenses.  We use the stacked images
from the final T0007 release taken from the Terapix
website\footnote{{http://terapix.iap.fr/cplt/T0006-doc.pdf}} for this
work.

We note that the \cfhtls is a niche survey with a unique combination of wide
imaging with deep sensitivity. It is a precursor to the ongoing wide imaging
surveys such as the Dark Energy Survey (DES), Kilo Degree Survey (KiDS) and the
Hyper Suprime-Cam (HSC) survey and other planned future surveys such as the
Large Synoptic Survey Telescope (LSST) survey.  The search for lenses with \sw
in the \cfhtls is an important step to learn lessons and prepare for lens
searches in these larger imaging surveys.

@@

The \cfhtls has been searched for lenses using various lens finding
methods and algorithms. Here, we give a brief summary of previously published lens
samples in the chronological order.

@@

\citet{Sygnet2010} carried out a search for edge-on galaxy lenses in the
\cfhtls WIDE. They identified galaxies, using {\sc Sextractor}, which
had $18<i<21$ and inclination angle $<25$~deg. After applying few more
selection criteria and visual inspection, they found about 3 promising
and a total of 18 lens candidates.

The \af \citep{More2012} was used for finding blue arc-like features in
the entire \cfhtls imaging without any pre-selection on the type of the lensing
object. The algorithm measures the second order moments of the flux distribution
in pixels within small regions to estimate the direction and extent of local
elongation of features. Pixels with high values of elongation are connected to
form an arc candidate. Finally, a set of thresholds on arc properties such as
the area, length, width, curvature and surface brightness are used to select
arc-like candidates. The search was carried out in the g-band which is the most
efficient wavelength to find typical lensed features.  This sample, called
SARCS, has 55 promising and a total of 127 lens candidates which are selected
from both \cfhtls WIDE and DEEP fields. The SARCS sample consists of some
galaxy-scale candidates and mostly groups/cluster scale lens candidates. This is
because more massive systems produce arcs or lensed images with large image
separation from the lensing galaxy which are easier to detect compared to the
galaxy-scales. In the absence of a large systematically followed up
verified sample of candidates, we choose the most promising 26 systems
as our bonafide lens sample from the \cfhtls WIDE. Also, the total
number of lens candidates is 108 from the \cfhtls WIDE alone.

In \citet{Elyiv2013}, the authors visually inspected a sample of 5500 optical
counterparts of X-ray point-like sources identified in the XMM-LSS
imaging of the \cfhtls W1 field. The goal was to find instances
of lensed quasars. Their sample consists of a total of 18 candidates out
of which 3 candidates are found to be promising.

The \rf \citep{Gavazzi2014} was used for finding compact rings or arcs
around centers of isolated and massive early-type galaxies. The \rf
subtracts the PSF-matched $i$-band images from the $g$-band images, and
looks for excess flux in the bluer $g$-band. An object detector measures
the properties of these residual blue features, and candidates which
meet the length-width ratio and tangential alignment criteria are then
visually inspected to form the final sample. \citet{Gavazzi2014}
pre-selected $\sim$638,000 targets as either photometrically-classified
early type galaxies, or objects selected to have red centers and blue
outer parts, from the T0006 CFHTLS data release catalogs. A total of
14370 galaxies were found to show detectable blue residuals, and 2524
were visually inspected, having passed the automatic feature selection
process. This led to a total of 330 lens candidates out of which 42 are
good quality (\texttt{q\_flag} = 3) and 288 are medium quality
(\texttt{q\_flag} = 2) candidates. In addition to the main well-defined
sample of \citet{Gavazzi2014}, a further 71 candidates were reported to
be detected by earlier version of the \rf or from the \cfhtls DEEP.
From the main sample of ``\rf candidates,'' the SL2S team found, during
their follow-up campaign, 33 confirmed lenses.
%(and 17 promising candidates) AM: PJM, what is this sample of 17 ?

The work by \citet{Maturi2014} used the arc finding code of
\citet{Seidel2007} and color properties of typical arcs to optimize arc
detection. This new approach was tested on the CFHTLS-Archive-Research
Survey \citep[CARS,][]{Erben2009} which covers an area of 37~sq.~deg
only and this entire data was also visually inspected by the authors to
estimate completeness and purity of their robotic search. They found 29
candidates with the robotic search alone and 41 candidates through pure
visual inspection some of which are known from previous searches. Most
of these candidates are medium-low
probability\footnote{http://www.ita.uni-heidelberg.de/˜maturi/Public/arcs}.

The \rf and the \af searches are the only searches that make use of a
lens finding algorithm and that has been run on the entire \cfhtls
imaging.  Thus, these can be considered as our reference sample of known
lenses from robotic searches. For the purposes of transparency and to
help with the training, the volunteers participating in \sw-\cfhtls lens
search were made aware of these two known lens samples. Images
containing the systems from the \rf and the \af samples were labelled as
``Known Lens Candidates'' in the \sw discussion forum,
TALK\footnote{http://talk.spacewarps.org/}, where volunteers have the
opportunity to discuss their findings with fellow volunteers and the
science team. In this paper, we refer to the sample of 330 \rf and 108
\af lens candidates as the sample of ``known lens candidates'' and the
sample of confirmed (or most promising) 33 \rf and 26 \af as the sample
of ``known lenses''. Note that the ``known lens'' sample is a subset of
the ``known lens candidates'' sample.

@@

In order to perform a blind lens search over the entire \cfhtls
\textsc{wide}, we present the volunteers with cutouts of images selected
from the survey region. We briefly describe the image presentation here
for completeness but more information can be found in Paper I. We use
the $g$, $r$ and $i$-band imaging from \cfhtls which are most useful for
visual identification of lenses.  We made color composite images using
the publicly-available code, \humvi\footnote{The open source  color image
composition code used in this work is available from
\texttt{http://github.com/drphilmarshall/HumVI}} following the
prescription of \citet{Lupton2004}. The color scales are chosen to
maximize the contrast between faint extended objects. These parameters
are then fixed during the production of all the tiles, in order to allow
straightforward comparison between one image and another, and for
intuition to be built up about the appearance of stars and galaxies
across the survey.

We extract contiguous cutouts of size 81.84\arcsec\ (440 pixels) with an
overlapping region of 10\arcsec (54 pixels) between the neighbouring
cutouts. This  results in a catalog of $\sim$ 430 000 cutouts for the entire \cfhtls
\textsc{wide} region. The size of the individual cutout is determined by
optimising factors such as the typical angular scales of gravitational
lenses, the number of objects seen in a single cutout and the total
number of image cutouts in the survey. If a lens
candidate happens to be too close to the edge of a cutout, then the overlap
between neighbouring cutout allows a volunteer to get a clearer view
of the same candidate in at least one of the cutouts. We note that since
the images are shown randomly, a volunteer may not necessarily come
across the neighbouring cutout unless the volunteer classifies a large
number of images. This is not a problem since our user base is extremely
large and we receive multiple classifications for the same cutout.
%We also provided access to the $u$ and $z$ band imaging
%data through dashboard, when the volunteers wished to take a closer look at a particular system\footnote{The dashboard tool page can be accessed from
%http://tools.zooniverse.org/\#/dashboards/spacewarp}.

@@

The simulated lenses are important to train citizen scientists who are
novice to the task of finding lenses but they are also crucial for
analyzing the classifications performed by the citizen scientists (more
details can be found in Paper I but see \Sref{sec:swap} for a brief
summary). In this section, we describe the framework used for generating the
simulated lens sample and give details of the sample itself along with
some known limitations.

@@

For the purpose of generating simulated lens systems, we divide them
in to two main categories a) galaxy scale lenses b) and group or
cluster-scale lenses. We further subdivide galaxy scale lenses based on the
type of the background sources, namely galaxies and quasars. We do not simulate
group scale - quasar lenses as they are expected to be even more rare. We now
describe our procedure to generate these different types of lens systems.

@@

In order to calculate the lensing cross-section, we first calculate the
luminosity of each potential lensing galaxy using the photometric redshifts
($z_l$) from the LRG catalog. We use the $L-\sigma$ scaling relation from the
bright sample of \citep{Parker2005} to set the velocity dispersion of the halo
hosting the galaxy which will be later used in the model
\be
\label{magstar2}
\sigma_v=142 \left(\frac{L}{L_{*}}\right)^{1/3} \,.
\ee
We assume that the knee of the luminosity function of galaxies, $L_*$ evolves
such that there is a decline of $1.5$ magnitudes between $z=1$ to $z=0$
\citep{Faber2007}.\footnote{We anchor our $L_*$ evolution at low redshifts using the
determination of $L_*$ in the r-band by \citet{Blanton2001}. To maintain
consistency in magnitude systems, we have converted the CFHT MegaCAM magnitudes
to SDSS magnitudes and k-corrected them to $z=0.1$.}

@@

For quasars, we assume the luminosity function prescription of \citep{Oguri2010}
and adopt k-corrections by \citep{Richards2006}.
%alp=-0.5;
%kcorr=-2.5*(1 + alp)*log10(1+zz);
%Dlum=cc.Dlofz(zz)/p.hval;
%DM=5*log10(Dlum);
%Mabs=mag-kcorr-DM-25.0;
The luminosity function is expressed as
\be
\frac{{\rm d}\Phi}{{\rm d}M}=\frac{\Phi_{*}}{10^{0.4(\alpha+1)(M_{\rm abs}-M_{*})} + 10^{0.4(\beta+1)(M_{\rm abs}-M_{*})} }
\ee
where the normalization, $\phi_{*}=5.34\times10^{-6} h^3$ Mpc$^{-3}$ and break
magnitude, $M_*=-20.90 + 5 {\rm log} h - 2.5 {\rm log} f(z)$. The redshift
dependent factor in $M_*$ is given by
\be
f(z)=\frac{e^{\zeta z_s}(1+e^{\xi z_*})}{(\sqrt{e^{\xi z_s}}+\sqrt{e^{\xi z_*}})^2} \,.
\ee
We adopt the best-fit values $\zeta=2.98$, $\xi=4.05$, $z_{*}=1.60$
\citep{Oguri2010}. For the faint end slope, we use $\beta=-1.45$ whereas for
the bright end slope, we use $\alpha=-3.31$ when $z_s<3$ and $\alpha=-2.58$ at
higher redshifts, as prescribed by \citep{Oguri2010}.
respectively.

@@

Next, we determine properties of the background source for every lens. We follow
similar procedures for both background galaxies and quasars. We perform
ray-tracing for all of the $N_{\rm src}$ sources using the publicly available
code \gravlens \citep{Keeton2000} and choose sources that satisfy our
selection criteria. We determine fluxes of the lensed images and the total
magnification of each of the lensed source. We draw a random source for which
the flux of the second brightest lensed image and the total magnification of all
lensed images are above the thresholds given in Table~\ref{tab:thresh}.

Since we want to produce realistic looking lens systems, we simulate lenses in
each of the five \cfhtls~filters. The colors of the background source galaxies are drawn
randomly from the photometric CFHTLenS catalog
\citep{Hildebrandt2012,Erben2013}.  Similarly, we use a quasar catalog from the
SDSS Data Release 9 \citep{Paris2012} from which colors are drawn to simulate
quasar lenses. Next, we assume deVaucoleur's profile to account for the size
and shape of the galaxies. The ellipticity and the position angle (PA) are drawn
randomly between the range given in Table~\ref{tab:thresh}. The effective
radius of the galaxy is estimated from the Luminosity$-$size relation
\citep{Bernardi2003} (with a redshift scaling to account for some size
evolution) given by
\be
R_{\rm eff}= 10^{0.52} \frac{L_r^{2/3}}{{(1+z_s)}^2}
\ee
where $L_r=L_s/10^{10.2}$. On the other hand, quasars are assumed to follow a
Gaussian profile where the $\sigma$ is equated to that of the median seeing for
every filter. The median seeing values are taken from Table 4 of the official
Terapix T0007 release explanatory document \footnote{
    http://terapix.iap.fr/cplt/T0007/doc/T0007-doc.pdf}.
